---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bebop Grilled Cheese 
### Dashboard Case Competition
Note: This project was undertaken in a team of five. The code and analysis contained in this repo is primarily my individual contribution; unless otherwise indicated it can be assumed to be my own. The contribution of my teammates has either been left out - as I do not have their code, or I have identified it. There were also analyses and charts that did not make it to the final product. As this was created for the final project of a course and a case competition, we had some restrictions. We were required to use R and Google Sheets for data processing, and our dashboard had to be developed using Google Sheets with four to five panels.

## Table of Contents

1. Case Description
2. 

## 1. Case Description
Bebop Grilled Cheese (BGC) is a food services business that operates four specialty food trucks, each with very similar menus across three product categories: main course, snack, and drink. As a specialty food truck operator, BGC is known for its gourmet grilled cheese sandwiches, special fried chicken sandwich, and veggie burgers.  All trucks also offer consumers hand cut truffle fries and a choice of different beverages.

* Each truck targets a different consumer market given its location so Alex adjusts menu prices slightly and adapts the truck’s operating schedules to fit the target market.
* The truck in the downtown metro area and the truck in the university district both operate year-round for six days a week.  
* The truck in the suburban area, which is located near parks and recreational centers, operates for a six-month period – May to October – for five days a week.
* The fourth truck operates for a six-month period – May to October – and moves to various locations around the city for special events, operating only on Fridays, Saturdays, and Sundays.
* Each truck operates with a minimum of two BGC employees and, during the busier months, adds a third employee for specific day parts – e.g., lunch time – when traffic is typically high.
* Chris sources all ingredients for the trucks so the average variable cost per product item for the 2022 fiscal year is the same across all four trucks.
The co-founders provided you with a copy of the operations report for the fiscal year-ended October 31, 2022. When BGC was founded, they set a target of 70% for gross margin because the food services company margin must cover labor costs for food truck employees as well as other fixed selling, general, and admin expenses.
**Business Problem:** BGC's gross margin in the 2022 fiscal year was below the target of 70%. What led to this result and how might we adjust operations to meet 70% gross margin next fiscal year?

## 2. Data
[Data set](https://github.com/AllyBMa/Portfolio/blob/main/Bebop%20Grilled%20Cheese/bebopSalesTransactions_2022.csv)

Contains 195,728 itemized sales transactions and the following 12 variables: FY, SalesDate, month, weekDay, truckID, orderID, lineItemID, group, item, price, q, cost

* FY = refers to the company’s fiscal year which runs from 2021-11-01 to 2022-10-31.
* SalesDate = sales date.
* month= month of sales date.
* weekDay = day of the week of sales date.
* truckID = truck ID takes four values: metroFY, specEventHY, suburbHY, uniDistFY.
* orderID = refers to the receipt number.
* lineItemID = refers to itemized ID of products shown in the receipt.
* group = refers to one of the three products groups: : Drink, Main, Snack.
* item = product name/description.
* price = product price.
* q = units sold.
* cost = cost of product (cost of good sold). *note: to simplify their accounting, and understanding that ASPE does not require presentation of cost of sales, BGC has opted not to classify any expenses as cost of sales. Expenses are either classified as COGS or operating expenses, based on their nature.

## 3. Data Processing and Analysis

```{r}
# attaching necessary packages
library(tidyverse)
# prevent scientific notation
options(scipen = 99)
# import and view data
dt0 <- read_csv("bebopSalesTransactions_2022.csv")
glimpse(dt0)
# check for missing data
dt0 %>% is.na() %>% colSums()
```

Perfect, no missing data. However, there are some variables that come in handy. Revenue is something that management wanted to examine, as well as a main input along with COGS for gross profit and gross profit margin.


```{r}
# new data set with revenue, cost of goods sold, 
# gross profit, and gross profit margin
dt1 <- dt0 %>% 
  mutate(
    revenue = price*q,
    cogs = cost*q,
    gp = revenue - cogs,
    gpm = gp/revenue
  )
glimpse(dt1) #view new data set
```

Now that we have our data with a more complete set of variables, let's pull out the things needed for each panel.

### Panel 1

Our first panel was to provide and overview of sales revenue, COGS, gross profit, and gross margin; and sales revenue and gross margin by month by truck/BGC, so we included three elements.

My teammates provided the overview in dollars for BGC, and the analysis by month by truck/BGC. I worked on revenue distribution - how much went to covering COGS and how much was profit.

```{r}
panelone <- dt1 %>% 
  select(
    month, truckID, revenue, cogs, gp, gpm
  ) %>% 
  group_by(month, truckID) %>% 
  summarize(SUMrev = sum(revenue),
            SUMcogs = sum(cogs),
            SUMgp = sum(gp),
            SUMgpm = sum(gpm)
            )
panelone #revenue, cogs, gp, and gpm of each truck by month
```

I then summarized this data in two pivot tables. Due to overlap with my teammate's chart, I ended up not using the grouping by month.

From these pivot tables we can see the dollar value of COGS and gross profit for each truck and BGC as a whole. For example, truck metroFY made a gross profit of \$251144.97, and spent \$110859.98 on COGS.

### Panel 2

Our second panel examined sales revenue for the year by truck, segmented by product item and grouped into the three product categories: main, snack, and drink. 

```{r}
paneltwo <- dt1 %>% 
  select(truckID, group, item, revenue)
paneltwo # revenue made by each item at each truck
```

After pulling out the relevant data, I continued data processing with pivot tables. 
We can see how much revenue each product earns at each truck/BGC. For BGC as a whole, Hand-Cut Truffle Fries earn $160745.39 of revenue.

In this table we can see what percentage of revenue each product group makes up for each truck/BGC. While Snack makes up ~15% for the other trucks, it makes up 31.16% for truck specEventHY. Main is the dominant product group for all trucks/BGC. 

### Panel 3

Panel three compared sales quantities and average sales prices by product item for BGC (total company) and for each of the trucks.

```{r}
dtprice <- dt1 %>%
  group_by(item)%>%
  summarize(itemavgprice=mean(price))
dtprice # average price of item across food trucks

dtq <-dt1 %>%
  group_by(group, item, truckID)%>%
  summarize(SUMq=sum(q))
dtq # number of each item sold at each truck

panelthree <- full_join(dtprice, dtq) #combine price and q
```
Note: There was an error in the case, it said prices of each item varied across the food trucks, but the prices in the data did not. As such price and avg price are the same, but we chose to follow the case and talk about avg price.

Once again, the data was processed in pivot tables - one for average price and one for quantity. Comparing the two, we see that as average price rises, quantity decreases.

### Panel 4

Data processed by teammate. Compares sales quantity of a specific truck to the average sales quantity of all trucks for each day of the week.

### Panel 5

The aim of panel five was to identify well performing product items and groups and areas of improvement. 

```{r}
dtproduct <- dt1 %>%
  group_by(item)%>%
  select(truckID, item, group, gpm)%>%
  distinct()
dtproduct # profit margin of each item

dtgroup <- dt1 %>% 
  group_by(group, truckID) %>% 
  select(truckID, group, gpm) %>% 
  summarize(groupPM = mean(gpm)) %>% 
  distinct()
dtgroup # profit margin of each group
```

One last time, I placed the data in pivot tables.

It can be seen that Coffee is the item with the highest profit margin of 83%; and Strawberry Banana Juice, Double Fried Chicken Sandwich, Vegan Veggie Burger, and Veggie Burger have the lowest profit margin of 61%.

From this pivot table, we see that the profit margin of product group Main is consistently under target. 

## 4. Data Visualization

Our dashboard prepared using Google Sheets and was built to follow [Tufte's Principles for Visualization](https://docs.google.com/presentation/d/11FFz8X6pH8nGrmIDJbvRfTDCdPK60Ii9RAuQmiFvBm4/edit#slide=id.g19a0f2ce771_0_272). Interactive functionality on Google Sheets is limited, but we used VLOOKUP and data validation to update the entire dashboard with a change to one cell, outlined in purple beside Truck Selection. Conditional formatting and data point formatting were used to keep a consistent colour and labelling scheme. 

### Panel 1

Two other elements of visualization created by teammates.
Used `=VLOOKUP(IF(Dashboard!$I$2='Data Validation'!$A$6,"Grand Total",Dashboard!$I$2),$I$1:$J$6,2,false)` and `=VLOOKUP(IF(Dashboard!$I$2='Data Validation'!$A$6,"Grand Total",Dashboard!$I$2),$I$8:$J$13,2,false)` to find respectively the COGS and gross profit values of the selected truck. Then created a pie chart from these two values to present the percentage of gross profit (gross margin) and cogs that make up revenue. 

A gauge would be a better visualization of gross margin, but gauge charts in Google Sheets are extremely limited and frankly don't look too great.

### Panel 2

**Element 1:** Used `=vlookup(If(Dashboard!$I$2="BGC (all)","Grand Total",Dashboard!$I$2),$A$4:$V$8,B11,false)` to find the revenue earned by each item of the truck selected. Created a column chart from these values.
**Element 2:** Used `=vlookup(If(Dashboard!$I$2="BGC (all)","Grand Total",Dashboard!$I$2),$A$23:$E$29,2,false)` to find the percentage of revenue that each product group made up for the selected truck. Used these values to create a donut chart.
**Element 3:** Used `=FILTER('Panel 2 Pivot'!$12:$12, 'Panel 2 Pivot'!$13:$13 = LARGE('Panel 2 Pivot'!$13:$13,1))`, with the last argument of `LARGE` being 1, 2, and 3 to display the name of the items that made the first, second, and third most revenue respectively for the selected truck.

### Panel 3

Used `=VLOOKUP(IF(Dashboard!$I$2='Data Validation'!$A$6,"Grand Total",Dashboard!$I$2),$A$3:$V$8,$B$17:$V$17,false)` and `=VLOOKUP(IF(Dashboard!$I$2='Data Validation'!$A$6,"Grand Total",Dashboard!$I$2),$A$12:$V$16,$B$17:$V$17,false)` to find the average price, and quantity bought of each item for the selected truck. Used these values to create a combo chart to visually compare the correlation between average price and quantity bought.

### Panel 4

Visualization created by teammate. Double line chart that compares sales quantity of the selected truck to the average sales quantity of all trucks each day of the week.

### Panel 5

**Element:** Used `=vlookup(If(Dashboard!$I$2="BGC (all)","Grand Total",Dashboard!$I$2),$G$1:$AB$8,H10,false)` to find the profit margin of each item for the selected truck. Created a column chart with scale up to 100%, essentially working as multiple gauges.
**Element:** Used `=vlookup(If(Dashboard!$I$2="BGC (all)","Grand Total",Dashboard!$I$2),$E$1:$I$7,2,false)` to find the profit margin of each group and truck specific menu for the selected truck. Created a column chart with scale up to 100%, essentially working as multiple gauges.

## 5. Key Points

Panel 1 (Teammate)
* Gross profit and revenue are highest in the months May to October, where all four trucks are running
* At 70.6%, the truck uniDistFY is the only one that meets the gross profit target of 70%
* Other trucks under target by 0.6% to 2.7% resulting in BGC's overall gross profit margin also just below target at 69.2

Panel 2
* Main product category makes up bulk of revenues, 61% for BGC as a whole and from 56.80% to 66.86% for each of the trucks
* Hand-Cut Truffle Fries and Honey Bacon on Triple Cheese Melt appear in the top three list of product items that generate the most revenue for every truck and the overall company

Panel 3 (Teammate)
* There is an inverse relationship between average item price and sales quantity. As the average item price rises, the quantity bought falls

Panel 4 (Teammate)
* Truck metroFY makes the highest quantity of sales on Friday
* Even though specEventHY only operates Friday to Sunday, on its operational days it still sells below or similar to average
* On average, trucks make the highest quantity of sales on Monday

Panel 5 
* Coffee is the item with highest profit margin, at 83.22%; followed by Redbull at 80.91%; and Polar Soda at 76.67%
* The items with the lowest profit margin of 61% are Strawberry Banana Juice, Double Fried Chicken Sandwich, Veggie Burger, and Vegan Veggie Burger
* Drink product group has highest profit margin for all trucks except for specEventHY

## 6. Recommendation
To increase BGC's gross profit margin and meet the targeted 70% in the next fiscal year, BGC Management should take the following two actions:
* Remove Strawberry Banana Juice and promote other drinks, or reduce the costs. It is only offered by truck suburbHY. Though it is a popular item, the most sold at that truck, it is one of the items with the lowest profit margin.
* Widen availability of Coffee. Currently only offered by truck metroFY, but has the highest profit margin of all BGC's offerings. 
